---
title: "Understanding Stock Price Beahavior using an R Based Analytical Framework"
author: "Francis Kurian"
date: "10/8/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
\section*{Introduction}
Stock market behavior is a well researched area with plenty of historic information available free. Several studies link the stock price movement of a company to the perception of the market participant about that company.IBM Watson APIs enable tracking of company specific news and social media interactions through extensive text mining and help come out with a Sentiment (Negative and positive emotions expressed through words or text) Index for a business.Business Sentiment Index(BSI) is one such index calculated by TRaiCE Fintech at company level. In addition, to understand  sensitivity of stock prices to Global events, Foreign Exchanges Rate is also introduced as an additional measure.  Idea here is to demonstrate how to extract various relevant data elements from diverse sources, transform and load them into an analytic framework to visualize and understand the inherent relationships.   

\section*{Objective}
Primary objective is to develop an analysis system using various R libraries. Extracting and processing data from multiple sources, data cleaning, simplifying repetitive tasks using control structures and functions, use of data visualization techniques and application of statistical methods  are the focus areas while building the framework in R. In other words, learn to write reproducible R code while analyzing stock price movement with respect to Business Sentiment Index(BSI) and Foreign Exchange Rate is the objective of this project.

\section*{List of Libraries}
For this project the following Libraries are used for data validation, graphics and statistical analysis

```{r libraries, echo=TRUE,message=FALSE,warning=FALSE,results=FALSE}
library(ggplot2) 
library(tidyr)
library(plyr)
library(lubridate)
library(scales)
library(reshape2)
library(summarytools) #dataframe summaries
library(ggfortify) #autoplot
library(sjPlot) #tabmodel
library(ggpubr) # wrapper for ggplot
```

\section*{Data sources, Extraction and Cleaning}
\subsection{Data files(.csv)used, Description and source}
### 1.List of companies to Analyze: Internally created CSV
### 2.Foreign Exchange daily data for various currencies: https://www.federalreserve.gov 
### 3.Daily Stock Prices: https://finance.yahoo.com 
### 4.Business Sentiments Index(BSI,IBM Watson API based): https://www.traice.io 

This project uses 6 different companies as test cases. However as long as the data files are available, companies can be added to the list and no code changes are necessary. Stock Price and BSI data will have 6 files each. Foreign Exchange is macro economic data so only one file with three different exchange rates are downloaded from the Federal Reserve. Daily data for two years(01-SEP-2019 to 31-AUG-2021) are used for the analysis.

\section*{Company Level Analysis -User Interface}
\subsection{To Analyze a company users can call this function with appropriate parameter changes. The graphs and regression results will be presented
for analysis and inference.
}
```{r csv1, echo=TRUE}
source("./src/fn_data_preperation.R")
fn.analysis("GE", "Stock.Price","bsi_score")
fn.analysis("F", "Stock.Price","USDxEUR")

```
\section*{Data Diagnostics-Understanding the input data better}
\subsection{Processing CSV files(Items 1-4 in Data sources, Extraction and Cleaning sections listed above)}
\subsubsection{Validates source csv files are existing for the companies and alerts the user otherwise.Process the files iteratively
Displays data summaries and charts to detect anomalies in the data}
```{r csv2, echo=TRUE}
fn.data_diagnostics()
```
\section*{R Code and Funtions used for analysis and diagnostics}
\subsection{fn.analysis }
```{r vis1, echo=TRUE}
fn.analysis <- function(company,var.Y,var.X){
  
CompanyNames <- './data/CompanyNames.csv'

# error trap and validation added - Peer review
  
  if (class(try(read.csv(CompanyNames, header = T),silent=T)) == "try-error") {
    print(message("Please change your Knit/Knit directory to 'Project Directory or 
    Current Working Directory'. This error normally appears when 
    R Markdown knit/Knit Direcotry is set to 'Document directory'"))
   break  } 
  else {
    df_companies <- read.csv(CompanyNames, header = T)
  }
  

  
  ForexFile <- './data/FederalReserve_CurrencyXchangeRate.csv'
  headers <- read.csv(ForexFile, skip = 5 , header = F, nrows = 1) #header at row #6
  df_forex <- read.csv(ForexFile, skip = 6, header = F) #data begins at row #7 
  colnames(df_forex) <- headers #apply headers 
  df_forex$period <- as.Date(df_forex$`Time Period`) #format as date
  
  df_forex$USDxINR <- as.numeric(gsub("ND","",df_forex$RXI_N.B.IN)) #clean & frmt as nmbr
  df_forex$USDxEUR <- as.numeric(gsub("ND","",df_forex$`RXI$US_N.B.EU`))
  df_forex$USDxYEN <- as.numeric(gsub("ND","",df_forex$RXI_N.B.JA))
  
  df_forex <- subset(df_forex,select = c(period, USDxINR, USDxEUR, USDxYEN))
  df_forex <- df_forex[!duplicated(df_forex$period),] # remove duplicates
  df_forex2 <- melt(df_forex, id.vars="period") # for all in one graphs
  
  sapply(df_forex, class)

  
  
  for (i in df_companies$ticker) {
    
    if (!file.exists(paste0("./data/",i, ".csv"))) {
      print("Please remove the Company. Files Missing for:")
      print(i)
      break
    }
    
    #process stock price/volume files from data directory
    #pick the csv file of every company in the input list and create a dataframe in a loop
    
    df_c <- read.csv(paste0("./data/",i, ".csv"), header = T) 
    df_c$period <- as.Date(df_c$Date) # add a date field 
    df_c$ticker <- i # attach ticker
    df_companies_subset <- subset(df_companies, ticker==i) 
    df_c <- merge(x=df_c, y=df_companies_subset, by="ticker",all.x=TRUE )#to get name
    assign(paste0("df_",i), df_c) #create a new data frame for later use
    
    # Appends company data frames to one for graphs.   
    
    if (exists("df_c_all")) {
      df_c_all <- rbind(df_c_all, df_c)
      
    }else {
      df_c_all=df_c}
    
    # Process Sentiments Index from data directory
    
    df_bsi <- read.csv(paste0("./data/",i, "_BSI.csv"), header = T) 
    df_bsi$period <- as.Date(parse_date_time(df_bsi$created_date,c('%Y-%m-%d','%m/%d/%y')))
    df_bsi$ticker <- i # attach ticker
    df_bsi <- merge(x=df_bsi, y=df_companies_subset, by="ticker",all.x=TRUE )##to get name
    df_bsi <- df_bsi[c('ticker','period','name','bsi_score')]
    
    assign(paste0("df_",i,"_BSI"), df_bsi) #create a new data frame forlater use
    
    # create a data frame to append  every data frame for graphs 
    
    if (exists("df_bsi_all")) {
      df_bsi_all <- rbind(df_bsi_all, df_bsi)
      
    }else {
      df_bsi_all <- df_bsi}
    
  }
  
  #### End of data extraction and cleaning process ####
  

 
 
 fn.regress <- function(company,var.Y,var.X){
   
   #create local dataframes based on the company names
   
   if (!exists(paste0("df_",company))) {
     print("Company Datafiles Missing.Process Terminated")
     return(FALSE)
   }else {
     
     df_b  <- get(paste0("df_",company,"_BSI"))
     df_s  <- get(paste0("df_",company))
     df_cname <- df_companies[df_companies[,1]  == company, ]
   }
   
   
   # check all three datasets for data issues and make it global
   
   d1 <- dfSummary(df_forex, style = "grid", plain.ascii = TRUE)
   # view(dfSummary(df_forex, style = "grid", plain.ascii = TRUE))
   d2 <- dfSummary(df_s, style = "grid", plain.ascii = TRUE)
   d3 <- dfSummary(df_b, style = "grid", plain.ascii = TRUE)
   
   df_forex <- na.omit(df_forex)
   d4 <- dfSummary(df_forex, style = "grid", plain.ascii = TRUE)
   
   # cat(df_cname[1,2])
   # print(d1)
   # print(d2)
   # print(d4)
   
   # merge all three datasets
   
   df_all <- merge(df_b, df_forex, by="period",all.x=TRUE )  # merge with Forex file
   
   df_all <- merge(x=df_all, y=df_s, by="period",all.x=TRUE )  # merge with stock price 
   
   df_all <- subset(na.omit(df_all),select=c(period,	ticker.x,	name.x,	bsi_score,	
                                    USDxINR,	USDxEUR,	USDxYEN,	Close,	Volume))
   
   # dfSummary(df_all, style = "grid", plain.ascii = TRUE)
   
   df_all <- rename(df_all, c("ticker.x"="ticker", "name.x"="CompanyName",
                              "Close"="Stock.Price", "Volume"="Stock.Volume"))
   
   d5 <- dfSummary(df_all, style = "grid", plain.ascii = TRUE) #output saved global
   
   # Scatterplot the variable relationship to visualize.Output saved global
   
   scatter <- ggplot(df_all, aes(x=df_all[,var.X], y=df_all[,var.Y])) + 
     geom_point()+
     geom_smooth(formula = y ~ x,method=lm)+ ggtitle(df_cname[1,2] ) +
     xlab(var.X) + ylab(var.Y)
   

   # Fit the simple regression model and create a summary
   
   
   multi.fit <- lm(Stock.Price~bsi_score+USDxEUR+USDxYEN+USDxINR, data=df_all)
   std_results <- summary(multi.fit) # std model results
   tab_results <- tab_model(multi.fit) # tabulated results
   residual_plot <- autoplot(multi.fit) # residuals plot
   
   print(scatter)
   cat("MULTIPLE REGRESSION ANALYSIS RESULTS")
   print(std_results)
   print(residual_plot)
   
   return(TRUE)
 }
 
 fn.regress(company, var.Y,var.X)
  
}
```
\section{User Defined Function for Company Level Analysis }
A generic function "fn.regres" was defined and it takes company name, dependent(x) and independent(y) variables for the regression analysis. It ensures that necessary data frames to conduct the analysis exists and polots, analysis summaries and other visual diagnostics are available in the results set.
```{r fn1, echo=TRUE}


```
\subsection{Executing the user defined Function to Test a Company with no data }
Expectation is that even if the user added a company name in to the list of companies to analyze and there is no data downloaded for that company,  this should not result in an error and the user should be alerted about missing data. 

```{r fn2, echo=TRUE}

```
\subsection{Executing the Analysis Function Repeatedly for various Companies in the list }
This outputs a scatterplot that shows the relationship between the variables, Standard Summary results of a regression function, Tabulated results of a regression function with coefficients and plots residuals for better diagnostics. Simple regression can be performed for any company and any set of variables. The coefficients and p values quantify the sensitivity of the variable and degree of confidence. Multiple regression and tests of time series properties are for the future development.   

```{r fn3, echo=TRUE}

```
\section{Summary of the Analysis }
Scatterplot and regression analysis demonstrate a significant relationship between Stock Prices and Business Sentiment Index and Stock Prices and USDvsEuro foreign exchange rates in case of General Electric Company. The same analysis can be repeated for any 6 companies in our test list by calling the function "fn_regression()" repeatedly with simply changing the function arguments.This is a rudimentary analysis with simple regression however significant effort was spent on reading and cleaning data  from various sources and made several data series consumable for various advanced algorithms that R offers. In the future, multiple regression  approach and time series models like co-integration can be explored  and that will add more analytic content to this project. 
\section{References}
1. Understanding Diagnostic Plots for Linear Regression Analysis. 
https://data.library.virginia.edu/diagnostic-plots
